PK object batch mode trial

### 2023-10-05 data-raw/batchMode_trial_gpm.Rmd
Major reorganization of vignette to specify how each figure in manuscript was made.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(DBI)
tidyverse_conflicts()
```

@@

1. Pulling Data from CvTdb  
  - Minimal `pk` object  
2. Running all possible fitting options  
  - Evaluating fitting options  
  - Head-to-head comparison of pooled vs joint models  
3. Running dose-normalized joint fits  
4. Analysis Plots  
  - CvTdb replicate variation (Figure 5)  
  - CvTdb replicate variation over ADME-normalized time (Figure 5B)  
  - CvTdb concentration and final time-point distributions (Supplementary Figure 1)  
  - invivoPKfit fit option selection (Figure 6)  
  - invivoPKfit model performance (Supplementary Figure 2)  
  - Model performance vs Data variability (Figure 7A)  
  - Goodness of fit plots (Figure 7B)  
  - Fits that may be improved (Supplementary Figure 3)  
5. Lombardo Analysis  
  - Chemicals included  
  - Comparison of derived TK stats  (Figure 8)
6. Benchmarking invivoPKfit
  - Parallelization speeds up the process of fitting (Supplementary Figure 4)

## Pulling data from CvTdb
This is largely taken from a `Querying CvTdb` vignette from Taylor Wall.

```{r query_functions}

postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
              user = Sys.getenv("CVT_USER"),
              password = Sys.getenv("CVT_PASS"),
              host = Sys.getenv("CVT_HOST"),
              dbname = Sys.getenv("CVT_DB_NAME"))
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) return(cat("\nMust provide a query to send"))
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN ", paste0("JOIN ", schema, "."), .))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

The following is the new SQL query, this will eventually be moved another vignette.

```{r SQL_queries}
# Sample query of CvTdb Documents SQL
# query <- "SELECT pmid, year, first_author, title FROM documents LIMIT 10"
# query <- "SELECT count(*) FROM studies"
# query <- "SELECT DISTINCT administration_route_normalized FROM studies"

# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  "a.id, a.time_original,  a.time_hr, a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",
  "b.id, b.analyte_name_original, b.analyte_dtxsid, b.analyte_casrn, ",
  "b.fk_analyzed_chemical_id, b.fk_test_chemical_id, ",
  "b.test_substance_dtxsid, b.analyte_name_secondary_original, ",
  "b.conc_medium_normalized, b.time_units_original, b.conc_units_original, ",
  "b.loq, b.loq_units, b.n_subjects_in_series, ",
  "c.id, c.administration_route_normalized, c.fk_dosed_chemical_id, c.dose_volume, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, c.dose_frequency, c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_original_units, ",
  "d.id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  "e.id, e.pmid, e.year, e.other_study_identifier, e.url, e.doi, e.extracted ",

  # Main Table
  "FROM conc_time_values a ",

  # Join with series table by series ID
  "LEFT JOIN series b ON b.id = a.fk_series_id ",

  # Join to studies table by study ID
  "LEFT JOIN studies c ON c.id = b.fk_study_id ",

  # Join to subjects table by subject ID
  "LEFT JOIN subjects d ON d.id = b.fk_subject_id ",

  # Join to documents table by extraction document ID
  "LEFT JOIN documents e ON e.id = c.fk_extraction_document_id ",

  # Filtering steps
  "WHERE b.conc_medium_normalized IN ('blood', 'plasma') AND ",
  "c.administration_route_normalized IN ('iv', 'oral')"
)

sql_query <- query_db(query = query,
         schema = "cvt")

names(sql_query)
# Need to change column names for this
{
  names(sql_query)[1] <- "conc_time_id"
  names(sql_query)[8] <- "series_id"
  names(sql_query)[22] <- "study_id"
  names(sql_query)[34] <- "subject_id"
  names(sql_query)[41] <- "document_id"
}

# Harmonize values for easier analysis
sql_query <- sql_query %>%
  mutate(
    species = tolower(species),
    sex = tolower(sex),
    loq = as.numeric(loq),
    n_subjects_in_series = as.numeric(n_subjects_in_series),
    dose_level_normalized = as.numeric(dose_level_normalized),
    dose_level_original = as.numeric(dose_level_original),
    dose_volume_units = str_extract(dose_volume,
                                    pattern = "(?<=[:space:]).+"),
    dose_volume = as.numeric(str_extract(dose_volume,
                                         pattern = "^(\\d|\\.)*")),
    dose_duration_units = str_extract(dose_duration,
                                      pattern = "(?<=[space]).+"),
    dose_duration = as.numeric(str_extract(dose_duration,
                                pattern = "^\\d"))
  )
sql_query <- sql_query %>%
  rename_with(~gsub(pattern = "fk_", replacement = "",
                         .x), .cols = contains("fk_"))

# Copying
cvt_df <- sql_query # Copy so I don't have to do query again to reset any errors

# Check other_study_identifier == S0916 so that it has dose_level_original_units == ug/mL
cvt_df %>% filter(other_study_identifier == "S0916") %>%
  pull(dose_level_original_units) %>% unique()
# There are still columns that are mg/kg... changing these to ug/mL
cvt_df <- cvt_df %>% mutate(
  dose_level_original_units = ifelse(
    other_study_identifier == "S0916",
    "ug/mL",
    dose_level_original_units
  )
)

# Need to convert the units to mg/kg
# In pulling_oral_iv_data_cvtdb.Rmd, it is noted that oral was 5mL/kg and iv 2mL/kg
# So I need to do ug/mL * mL/kg * 1mg/1000ug = mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0916",
      ifelse(administration_route_normalized %in% "oral",
             dose_level_original * 5/1000,
             dose_level_original * 2/1000),
      dose_level_normalized),
    dose_level_corrected_units = "mg/kg",
    .after = dose_level_original_units)

# For this study, it was in ng/kg not mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0624",
      dose_level_original * 1E-6,
      dose_level_normalized),
    .after = dose_level_original_units)

# The following measured C-14/H-3 but are indicators for the following Chemicals
# document_id = 65, DTXSID = DTXSID7031290, name = Isocyclemone E or beta-OTNE
# document_id = 61, DTXSID = DTXSID4022527, DTXSID3020209,
### name(s) = propylparaben, butylparaben, missing...methylparaben
# Propylparaben SERIES_ID = 110, Butylparaben SERIES_ID = 112
# For above, HPLC reviealed a single peak corresponding with p-hydroxybenzoic acid (PHBA, DTXSID3026647)
# document_id = 46, DTXSID = DTXSID2020139, name = benzo[a]pyrene
# document_id = 31, DTXSID = DTXSID2020139, name = benzo[a]pyrene

# However, now would be good time to eliminate any ones where they analyte != test chemical
# Some studies may have dosed with multiple chemicals... important to note
# cvt_df <- cvt_df %>% filter(analyzed_chemical_id == test_chemical_id)
# For now, I will change the chemical name and dtxsid
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 65, "DTXSID7031290",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 65, "beta-OTNE",
                                        analyte_name_original)) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "DTXSID4022527",
                                   series_id %in% 112 ~ "DTXSID3020209",
                                   .default = analyte_dtxsid
                                 ),
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "Propylparaben",
                                   series_id %in% 112 ~ "Butylparaben",
                                   .default = analyte_name_original
                                 ),
                                 analyte_name_original)
         ) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% c(46, 31),
                                 "DTXSID2020139",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% c(46, 31),
                                        "benzo[a]pyrene",
                                        analyte_name_original))

# For study S0976, Caffeine, L-ephedrine, and psuedoephedrine were dosed
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    other_study_identifier %in% "S0976",
    case_when(
      analyte_name_original %in% "Caffeine" ~ 4.00,
      analyte_name_original %in% "L-Ephedrine" ~ 4.44,
      analyte_name_original %in% "Pseudoephedrine" ~ 4.43
    ),
    loq),
    loq_units = ifelse(
      other_study_identifier %in% "S0976",
      "ng/mL",
      loq_units
    ))

# Some pentachloranisole studies
# S0361 seems to have some data that has not been extracted (also for mouse)
cvt_df <- cvt_df %>%
  mutate(loq = case_when(
    other_study_identifier %in% "S0361" ~ 0.05,
    other_study_identifier %in% "S0328" ~ 1, # This is inferred
    # Highest value under loq is 0.9
    # Lowest value above loq is 1.263, but there are differences in GC-EC and HPLC methods
    .default = loq
  ),
  loq_units = case_when(
    other_study_identifier %in% "S0361" ~ "ug/mL",
    other_study_identifier %in% "S0328" ~ "ug/mL",
    .default = loq_units
  ))

# Checking if entire groups of data are NA for conc and loq
# Really only eliminates Tamoxifen
# cvt_df %>%
#   # filter(analyte_dtxsid %in% "DTXSID1034187") %>%
#   group_by(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>%
#   filter(all(is.na(conc) & is.na(loq))) %>% 
#   distinct(analyte_dtxsid, document_id, analyte_name_original,
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>% View()

# Checking individual chemicals
# cvt_df %>%
#   filter(analyte_dtxsid %in% "DTXSID2020682") %>%
#   distinct(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected)

# PMID 6310077 is a tritium tracing of B[a]P and these were given as % of Dose
# The initial dose was 117 nmol, 2 mCi/kg
# Molecular Mass of benzo[a]pyrene (C20H12) is 252.316 g/mol
# The initial dose *I think* is in dose_level_normalized
cvt_df <- cvt_df %>%
  mutate(conc = ifelse(document_id %in% 38,
                       conc_original*dose_level_corrected,
                       conc))

# NTP-CEB C96019 just has a lot of data below LOQ
# For S0916, there is no > LOQ data for anything but Tamoxifen
# N-di-butyl phthalate is only good at the iv dose
# DTXSID2021315 is only good at 1E-4 Dose
# S0623 Doesn't have any plasma data
# There is a maybe control sample for hexachlorobenzene (Dose = 0.0 ) ?

# Filter by whether at each dose, there are no LOQ or Concentration value reported
# This is the threshold for Minimally Informative Data in our use case.
cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq))) %>%
  distinct(analyte_dtxsid, species,) %>% nrow() # Says there are 433 distinct

cvt_df <- cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq)))

# Give N-desmethyltamoxifen it's due DTXSID501315924
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "N-desmethyltamoxifen",
    "DTXSID501315924",
    analyte_dtxsid
  ))

cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "Methemoglobin",
    "DTXSID301056085",
    analyte_dtxsid
  ))

# Standardizing the names (example: glucuronide had two entries)
cvt_df <- cvt_df %>%
  mutate(analyte_name_original = tolower(analyte_name_original))

# Eliminate study that gave chemical each day, not single dose
# This may be used in another type of model, but right now we deal with single bolus
# Only include those that have NA or 1 for dose frequency
cvt_df <- cvt_df %>%
  filter(dose_frequency %in% c("1", NA))

# Normalize LOQ units
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    loq_units %in% c("ug/L", "ng/mL"), loq * 1E-3, loq)) # mg/L and ug/mL are same scale
# Normalize conc_sd and conc units manually
cvt_df <- cvt_df %>%
  mutate(conc_sd_normalized = case_when(
    conc_units_original %in% c("ug/L", "ng/mL", "ng-Eq/mL", "ng/g tissue") ~ conc_sd_original*1E-3,
    conc_units_original %in% c("g/dL") ~ conc_sd_original*1E-4,
    conc_units_original %in% c("pg/mL") ~ conc_sd_original*1E-6,
    .default = conc_sd_original
  ))
# Need to normalize the ng-eq/mL sample, conc is generally already normalized
cvt_df <- cvt_df %>%
  mutate(conc = case_when(
    conc_units_original %in% c("ng-eq/mL") ~ conc_original*1E-3,
    .default = conc
  ))

cvt_df <- cvt_df %>%
  dplyr::filter(!(analyte_dtxsid %in% c("DTXSID501315924",
                                      "DTXSID10169854",
                                      "DTXSID2021103") )) # These only have 1 or 2 values

# More notes:
# DTXSID6020438 & DTXSID0020868 rat is from water (top) and oil (bottom curve) gavage

@@

### PK OBJECT ###----
# Minimum pk object, add options later
minimal_pk <- pk(data = cvt_df,
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
```

@@

### Evaluations-----
my_pk_000p <- do_fit(my_pk_000p, n_cores = 14) 
my_pk_000p <- my_pk_000p %>% evaluate_choices() # Done
my_pk_100p <- do_fit(my_pk_100p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_110p <- do_fit(my_pk_110p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_111p <- do_fit(my_pk_111p, n_cores = 8) %>% evaluate_choices() # Done
rm(my_pk_100p,my_pk_110p,my_pk_111p)
gc()
my_pk_010p <- do_fit(my_pk_010p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_011p <- do_fit(my_pk_011p, n_cores = 8) %>% evaluate_choices() # Done 
my_pk_001p <- do_fit(my_pk_001p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_101p <- do_fit(my_pk_101p, n_cores = 8) %>% evaluate_choices() # Done

@@

Dose-normalization and a hierarchical model were the options that produced the least
amount of error in predictions and derived TK stats.

```{r Standard_Settings}
# Full option setting
my_pk <- pk(data = cvt_df,
            mapping = ggplot2::aes(
              Chemical = analyte_dtxsid,
              Chemical_Name = analyte_name_original,
              DTXSID = analyte_dtxsid,
              CASRN = analyte_casrn,
              Species = species,
              Reference = document_id,
              Media = conc_medium_normalized,
              Route = administration_route_normalized,
              Dose = dose_level_normalized,
              Dose.Units = "mg/kg",
              Subject_ID = subject_id,
              Series_ID = series_id,
              Study_ID = study_id,
              ConcTime_ID = conc_time_id,
              N_Subjects = n_subjects_in_series,
              Weight = weight_kg,
              Weight.Units = "kg",
              Time = time_hr,
              Time.Units = "hours",
              Value = conc,
              Value.Units = "mg/L",
              Value_SD = conc_sd_normalized,
              LOQ = loq
            )) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
my_pk <- do_preprocess(my_pk)
my_data <- get_data(my_pk)

An aside: working with one DTXSID + Species combination

The first step is understanding how to work with a single DTXSID + Species combination.
Once we know how to do this, it will be easier to apply it to (potentially)
a list of these objects in a vectorized fashion.
Here I subset the `cvt` object and use that as my working data.

```{r data_loading}
# There are 103 unique chemicals in this object
cvt %>% pull(chemicals_analyzed.dsstox_substance_id) %>%
  unique() %>%
  length()

# easy fit
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID5020732")

Next, I want to follow the steps required for the particular analysis I would like.
In this instance, I set the optimizer to "L-BFGS-B" and dose-normalize the output.
I do not scale the time or log-transform the concentration values.
Since I would like to do a "Pooled" analysis,
I set the `error_group = vars(Chemical, Species)`.

```{r single_fit}
my_pk_original <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
# Add Reference for Joint Analysis

my_pk_original <- do_preprocess(my_pk_original)

@@

my_pk <- do_prefit(my_pk) # needs reframe somewhere
# Timing for normal fit sapply
my_pk <- do_fit(my_pk, 
                n_cores = 12)
my_pk <- do_fit(my_pk)

@@

### Figure 5A: Replicated timepoint variation in CvTdb

### Figure 5B: Variation over time in CvTdb data

### Supplementary Figure 1: Concentration and final time-points have wide range

### Figure 6: Selecting best fitting options for CvTdb data

### Figure 7A: Model performance vs Data Variability

### Figure 7B: Multiple goodness-of-fit metrics validate model performance

### Supp. Fig. 3: Fits that may be improved with other settings

### Figure 8: Comparing derived TK stats with human TK stats from Lombardo et al.

### Supp. Fig. 4: Parallelization decreases runtime of invivoPKfit

```{r cvt_analysis}
my_df %>%
  ggplot(aes(x = maxT/(24)
             )) +
  geom_histogram(fill = "grey5") +
  scale_x_log10() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  expand_limits(y = 0) +
  labs(x = "Day of final Timepoint",
       y = "Count") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(face = "bold"))

```{r allDTXSIDs_analysis}

# after fitting pk object for all cvt objects
# saveRDS(my_pk_100p, file = "data-raw/all_cvt_pk.rds")
# saveRDS(my_pk, file = "data-raw/all_cvt_pk_JOINT.rds")
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
my_coefs <- coef(my_pk, drop_sigma = FALSE)
get_winning_model(my_pk) %>%
  filter(model %in% c("model_2comp"),
         method == "L-BFGS-B") %>%
  dplyr::select(Chemical, Species) %>% distinct() %>% nrow()
all_my_data  %>% dplyr::select(Chemical, Species) %>% distinct() %>% nrow()
my_AIC <- AIC(my_pk)
my_tkstats <- get_tkstats(my_pk)
my_IC <- left_join(AIC(my_pk), BIC(my_pk))

my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "bobyqa") %>%
  ungroup() %>%
  count(model)
# model_1comp 59
# model_2comp 50
# model_flat   2

my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  # filter(model == "model_flat") %>% pull(Chemical) %>% unique()
  count(model)
# model_1comp 57
# model_2comp 50
# model_flat   4

my_preds <- predict(my_pk,
                    use_scale_conc = TRUE)

winmodels <- my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  dplyr::select(Chemical, Species,
                method, model)

my_preds <- inner_join(winmodels, my_preds)
my_tkstats <- inner_join(winmodels, my_tkstats) %>%
  dplyr::select(-TKstats) %>% distinct()
my_nca <- get_nca(my_pk,
                  nca_group = vars(Chemical, Species,))
my_nca <- my_nca %>% dplyr::select(-param_sd_z, -param_units) %>%
  pivot_wider(names_from = param_name,
              values_from = param_value)

all_my_data <- get_data(my_pk)
all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  filter(Route == "iv") %>%
  count(foldConc < 0.5) %>%
  {.[[2,2]]/(.[[1,2]] + .[[2,2]])}
# 88.9% inside twofold ## New CvT extraction: 91.3%
# 2.2%  above twofold  ## 1.7%
# 8.8%  below twofold  ## 7.0%
# 
#

all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  ggplot(aes(x = foldConc)) +
  geom_histogram(binwidth = 0.1,
                 fill = "grey5") +
  coord_cartesian(xlim =  c(0, 4),
                  ylim = c(0, 2000)) +
  scale_y_continuous(expand = c(0, 0)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  # annotate(geom= "text",
  #          x = 1.25, y = 1800,
  #          label = "88.9% of Observations\n within a factor of 2") +
  theme_classic() +
  labs(x = "Concentration Normalized to Average",
                y = "Number of Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedConcentrations_2023-09-21.png"),
       height = 3.5,
       width = 5.5,
       units = "in")

all_my_data %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         exclude == FALSE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  ggplot(aes(
    x = normTime,
    y = foldConc
  )) +
  geom_point(alpha = 0.1, size = 0.7) +
  geom_vline(xintercept = 1, color = "blue",
             size = 1, linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4",
             size = 1, linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2), color = "red3",
             size = 1, linetype = "dashed") +
  facet_grid(cols = vars(Route),
             rows = vars(Species),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Mean-Normalized Concentration") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ADME_NormalizedTime-NormalizedConcentrations_onlyDetects_2023-09-21.png"),
       height = 4,
       width = 6,
       units = "in",
       dpi = 300)

# Per chemical + Species average mean normalized concentration variability over time
all_my_data %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc,
         exclude == FALSE) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical, Species, Route, Media, Dose) %>%
  summarize(sdFC = sd(foldConc)) %>%
  mutate(maxDose = max(Dose),
         minDose = min(Dose)) %>%
  ungroup() %>%
  add_count(Dose, name = "Unique_Doses") %>%
  filter(Unique_Doses == 1) %>%
  arrange(desc(sdFC))

my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media, model, method,
                            tmax) %>%
              filter(!is.na(tmax))

all_my_data %>%
              dplyr::select(Chemical, Species,
                            Route, Media, Dose,
                            Reference) %>%
  distinct() %>% nrow()

my_preds <- predict(my_pk,
                    use_scale_conc = TRUE)

winmodels <- get_winning_model(my_pk)

my_preds <- inner_join(winmodels, my_preds)
my_tkstats <- inner_join(winmodels, my_tkstats) %>%
  dplyr::select(-TKstats) %>% distinct()
my_nca <- get_nca(my_pk,
                  nca_group = vars(Chemical, Species,))
my_nca <- my_nca %>% dplyr::select(-param_sd_z, -param_units) %>%
  pivot_wider(names_from = param_name,
              values_from = param_value)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(cowplot)
library(RColorBrewer)
tidyverse_conflicts()
# Note I have now saved the RDS file for JOINT model with L-BFGS-B
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

@@

1. Pulling Data from CvTdb  
  - Minimal `pk` object  
2. Running all possible fitting options  
  - Evaluating fitting options  
  - Head-to-head comparison of pooled vs joint models  
3. Running dose-normalized joint fits  
4. Analysis Plots  
  - CvTdb replicate variation (Figure 4A)  
  - CvTdb replicate variation over ADME-normalized time (Figure 4B)  
  - CvTdb concentration and final time-point distributions (Supplementary Figure 1)  
  - invivoPKfit fit option selection (Supplementary Table 1)  
  - invivoPKfit model performance (Supplementary Figure 2)  
  - Model performance vs Data variability (Figure 5)  
  - Goodness of fit plots (Figure 6)  
  - Fits that may be improved (Supplementary Figure 3)  
5. Lombardo Analysis  
  - Chemicals included  
  - Comparison of derived TK stats  (Figure 7A)  
  - Histograms for Steady-state Volume of Dist. and Total Clearance (Supplemental Figure 4)
6. Benchmarking invivoPKfit
  - Parallelization speeds up the process of fitting (Supplementary Figure 5)

## Pulling data from CvTdb

See `pulling_iv_oral_cvtdb.Rmd`.

@@

### Minimal PK Object ###----
# Minimum pk object, add options later
# Note that these mappings are now a default mappings, just being verbose here.
minimal_pk <- pk(data = cvt_df,
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
```

@@

### Evaluations-----
my_pk_000p <- do_fit(my_pk_000p, n_cores = 14)  %>% evaluate_choices() # Done
my_pk_100p <- do_fit(my_pk_100p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_110p <- do_fit(my_pk_110p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_111p <- do_fit(my_pk_111p, n_cores = 8) %>% evaluate_choices() # Done
rm(my_pk_100p,my_pk_110p,my_pk_111p)
gc()
my_pk_010p <- do_fit(my_pk_010p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_011p <- do_fit(my_pk_011p, n_cores = 8) %>% evaluate_choices() # Done 
my_pk_001p <- do_fit(my_pk_001p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_101p <- do_fit(my_pk_101p, n_cores = 8) %>% evaluate_choices() # Done

@@

Dose-normalization and a hierarchical model were the options that produced the least
amount of error in predictions and derived TK stats.
In addition, _L-GFGS-B_ was the optimizer of choice

```{r Final_Options}
my_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk <- do_preprocess(my_pk)
cvt_DTXSIDs <- unique(my_pk$data$Chemical)

@@

my_pk <- do_prefit(my_pk) # needs reframe somewhere

{
  start <- Sys.time()
  my_pk <- do_fit(my_pk, 
                  n_cores = 12)
  round(Sys.time() - start, 4)
}

@@

```{r data_aggregation}
# Need the following data for any of the subsequent plots

# after fitting pk object for all cvt objects or reading the fitted pk object in `setup`
winmodel <- get_winning_model(my_pk)
my_preds <- inner_join(predict(my_pk, use_scale_conc = TRUE), winmodel)
my_residuals <- residuals(my_pk, use_scale_conc = FALSE) %>%
  inner_join(winmodel)
my_tkstats <- eval_tkstats(my_pk)
my_nca <- get_nca(my_pk)
all_my_data <- get_data(my_pk)

fe_df <- fold_errors(my_pk, method = "L-BFGS-B") %>% inner_join(winmodel)
# May need to change the use_scale_conc defaults
r2_df <- rsq(my_pk, method = "L-BFGS-B", use_scale_conc = FALSE) %>%
  inner_join(winmodel)
myAIC <- AIC(my_pk, method = "L-BFGS-B") %>%
  inner_join(winmodel)
myRMSE <- rmse(my_pk, method = "L-BFGS-B",
               rmse_group = vars(Route, Media, Dose),
               use_scale_conc = FALSE) %>% 
  inner_join(winmodel)
